---
title: "Prey Delivery Analysis"
author: "Eliza Stein"
date: "11/8/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


# Libraries

```{r}

#install.packages("tidyverse")

library(tidyverse)


```



# Load data

```{r}

setwd("C:/Users/eliza/Documents/final-project")
pdOriginal <- read.csv("./data/pd_main.csv")


#rename the first column, which imported with a special character
names(pdOriginal)[1] <- "nest"


```




## Clean up data


Filter to only include males in B (MGSA) and C (HFSA).

```{r}

pdHM <- pdOriginal %>%
  separate(col = nest, into = c("study_site", "territory"), sep = 1, remove = TRUE) %>%
  filter(study_site == "B" | study_site == "C", sex == "M")

```


Check structure of data.

```{r, structure}

#str(pdHM) #columns 't180' and 't225' are character classes, "nest_age" is character class

#unique(pdHM$t180) #at least one cell has an asterisk after the value
#unique(pdHM$t225) #same here
#unique(pdHM$nest_age) #"pred" and "" can be converted to NA

#remove asterisks
pdClean <- pdHM %>%
  mutate(t180 = gsub("\\*", "", t180)) %>%
  mutate(t225 = gsub("\\*", "", t225))


#change these columns to numeric
pdClean$t180 <- as.integer(pdClean$t180)
pdClean$t225 <- as.integer(pdClean$t225)
pdClean$nest_age <- as.integer(pdClean$nest_age) #warning here is ok--NAs are replacing "pred" and "" values


```



# Prey Deliveries Throughout Night

Data was separated by study site (HFSA vs. MGSA) and by incubation vs. nestling stage. Nestling period is defined as nest_age >= 22 days. If nest age was not indicated in original dataset, field notes were used to determine whether nest was in incubation (all eggs) or nestling (at least one nestling) stage. For these records, the following values were manually input: nest_age = 0 for incubation or nest_age = 100 for nestling, so that this data could be easily separated from known nest age.  If it was later determined that nest had been predated before observation, "pred" was entered. If the nest stage could not be determined, it was left blank."pred" and "" values were converted to NA earlier when this column was converted to numeric.

since the last observation for HFSA is at time = 240, we'll end both datasets there.

First, create data frame with PDs for whole dataset, then broken down by each study site and stage.

```{r, pd_indiv}

#create independet dfs for HFSA (nestling and incubation stage) and MGSA (nestling and incubation state). 

#select relevant columns, add column for stage, rename study sites, drop NAs
pdStage <- pdClean %>%
  select(study_site, nest_age, t15:t240) %>%
  mutate(
    stage =
      ifelse(nest_age < 22, "incubation", "nestling"),
    study_site =
      ifelse(study_site == "B", "MGSA", "HFSA")) %>%
  drop_na(stage)

#change column names to remove "t" in front of time interval
colnames(pdStage) <- c("study_site", "nest_age", "15", "30", "45", "60", "75", "90", "105", "120", "135", "150", "165", "180", "195", "210", "225", "240", "stage")

#create independent dfs for each study site and stage
pdHFSA_inc <- pdStage %>%
  filter(study_site =="HFSA", stage == "incubation") %>%
  select('15':'240')

pdHFSA_nest <- pdStage %>% 
  filter(study_site =="HFSA", stage == "nestling") %>%
  select('15':'240')

pdMGSA_inc <- pdStage %>%
  filter(study_site =="MGSA", stage == "incubation") %>%
  select('15':'240')
         
pdMGSA_nest <- pdStage %>%
  filter(study_site =="MGSA", stage == "nestling") %>%
  select('15':'240')


```




## Mean PD tables

Four stand-alone data.frames, one for (incubation), one for HFSA (nestling), one for MGSA(incubation), and one for MGSA (nestling):

```{r}

meanHFSA_inc <- data.frame(
  time = as.numeric(colnames(pdHFSA_inc)),
  HGSA_incubation = colMeans(pdHFSA_inc, na.rm = TRUE))

meanHFSA_nest <- data.frame(
  time = as.numeric(colnames(pdHFSA_nest)),
  HGSA_nestling = colMeans(pdHFSA_nest, na.rm = TRUE))

meanMGSA_inc <- data.frame(
  time = as.numeric(colnames(pdMGSA_inc)),
  MGSA_incubation = colMeans(pdMGSA_inc, na.rm = TRUE))

meanMGSA_nest <- data.frame(
  time = as.numeric(colnames(pdMGSA_nest)),
  MGSA_nestling = colMeans(pdMGSA_nest, na.rm = TRUE))

```


Two tables, one for nestling stage and one for incubation stage. This is just for visualization, not used in further analysis.

```{r eval=FALSE, include=FALSE}

#filtered this one to end at time = 180, because no observations exist for MGSA after that point
meanInc <- filter(
  data.frame(
    time = as.numeric(colnames(pdHFSA_inc)),
    HGSA_incubation = colMeans(pdHFSA_inc, na.rm = TRUE),
    MGSA_incubation = colMeans(pdMGSA_inc, na.rm = TRUE)),
  time <= 180
)

meanNest <-
  data.frame(
  time = as.numeric(colnames(pdHFSA_nest)),
  HGSA_nestling = colMeans(pdHFSA_nest, na.rm = TRUE),
  MGSA_nestling = colMeans(pdMGSA_nest, na.rm = TRUE)
)

```


One table showing mean PDs for all study sites and stages:
```{r}
meanAll <- data.frame(
  time = as.numeric(colnames(pdHFSA_inc)),
  HGSA_incubation = colMeans(pdHFSA_inc, na.rm = TRUE),
  HGSA_nestling = colMeans(pdHFSA_nest, na.rm = TRUE),
  MGSA_incubation = colMeans(pdMGSA_inc, na.rm = TRUE),
  MGSA_nestling = colMeans(pdMGSA_nest, na.rm = TRUE)
)

```




## Calculate confidence intervals

Write and apply function to obtain CIs.

```{r}

#function to grab CIs from any vector
getCI <- function(vec, n_samp=10000) {
  smp <- replicate(n_samp, mean(sample(vec, replace = TRUE), na.rm = TRUE))
  CIs <-quantile(smp, c(0.025, 0.975), na.rm = T)
  return(CIs)
}

#now apply it across the columns of PD data
ciHFSA_inc <- apply(pdHFSA_inc, 2, FUN = getCI)
ciHFSA_nest <- apply(pdHFSA_nest, 2, FUN = getCI)
ciMGSA_inc <- apply(pdMGSA_inc, 2, FUN = getCI)
ciMGSA_nest <- apply(pdMGSA_nest, 2, FUN = getCI)
```


Add CIs to a data frame. Remove rows where time > 180 for ciInc because no data is available for MGSA after this time.

```{r}

ciInc <- filter(
  data.frame(
    study_area = c(rep("HFSA", nrow(meanHFSA_inc)), rep("MGSA", nrow(meanMGSA_inc))),
    mean = c(colMeans(pdHFSA_inc, na.rm = TRUE), colMeans(pdMGSA_inc, na.rm = TRUE)),
    ci_l = c(ciHFSA_inc[1,], ciMGSA_inc[1,]),
    ci_h = c(ciHFSA_inc[2,], ciMGSA_inc[2,]),
    time = c(as.numeric(rownames(meanHFSA_inc)), as.numeric(rownames(meanMGSA_inc))),
    stage = "Incubation"),
  time <= 180)


ciNest <- data.frame(
    study_area = c(rep("HFSA", nrow(meanHFSA_nest)), rep("MGSA", nrow(meanMGSA_nest))),
    mean = c(colMeans(pdHFSA_nest, na.rm = TRUE), colMeans(pdMGSA_nest, na.rm = TRUE)),
    ci_l = c(ciHFSA_nest[1,], ciMGSA_nest[1,]),
    ci_h = c(ciHFSA_nest[2,], ciMGSA_nest[2,]),
    time = c(as.numeric(rownames(meanHFSA_nest)), as.numeric(rownames(meanMGSA_nest))),
    stage = "Nestling")
    
ciAll <- rbind(ciInc, ciNest)

```



## Plot

Incubation:
```{r}

plotInc <- ggplot(data = ciInc) +
  geom_point(aes(x = time, y = mean, color = study_area, group = study_area),
             position = position_dodge(width=0.75)) +
  geom_errorbar(aes(x=time, ymax = ci_h, ymin=ci_l, color = study_area, 
                    group = study_area),
                position = position_dodge(width=0.75)) +
  labs(x = "Time After Sunset (minutes)", y = "Mean Prey Deliveries", 
       title = "Incubation", color = "Study Area") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```


Nestling:
```{r}

plotNest <- ggplot(data = ciNest) +
  geom_point(aes(x = time, y = mean, color = study_area, group = study_area),
             position = position_dodge(width=0.75)) +
  geom_errorbar(aes(x=time, ymax = ci_h, ymin=ci_l, color = study_area, 
                    group = study_area),
                position = position_dodge(width=0.75)) +
  labs(x = "Time After Sunset (minutes)", y = "Mean Prey Deliveries", 
       title = "Nestling", color = "Study Area") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

Both:
```{r}

plotAll <- ggplot(data = ciAll) +
  geom_point(aes(x = time, y = mean, color = study_area, group = study_area),
             position = position_dodge(width=0.75)) +
  geom_errorbar(aes(x=time, ymax = ci_h, ymin=ci_l, color = study_area, 
                    group = study_area),
                position = position_dodge(width=0.75)) +
  labs(x = "Time After Sunset (minutes)", y = "Mean Prey Deliveries", 
       title = "Average Prey Deliveries Throughout Night", color = "Study Area") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_grid(~ stage)

```

Plot both as separate plots, side by side:
```{r}

grid.arrange(plotNest, plotInc)

```



## Prey Deliveries by Year

Going to create a similar graph but with years on the x-axis. I will sum values from time = 15 through time = 60 and only use this data from the first hour of each night.

```{r}


```

