---
title: "Prey Delivery Analysis"
author: "Eliza Stein"
date: "11/8/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
break down by nestling/incubation
for Hayman nestling and incubation: dot and whisker plot, if pd changes over years

# Libraries

```{r}

#install.packages("tidyverse")
library(tidyverse)

```



# Load data

```{r}

setwd("C:/Users/eliza/Documents/final-project")
pdOriginal <- read.csv("./data/pd_main.csv")


#rename the first column, which imported with a special character
names(pdOriginal)[1] <- "nest"


```



# Tidy data

Filter to only include males in B (Missouri) and C (Hayman)

```{r}

pdHM <- pdOriginal %>%
  separate(col = nest, into = c("study_site", "territory"), sep = 1, remove = TRUE) %>%
  filter(study_site == "B" | study_site == "C", sex == "M")

```


Check structure of data

```{r}

str(pdHM) #columns 't180' and 't225' are character classes

unique(pdHM$t180) #at least one cell has an asterisk after the value
unique(pdHM$t225) #same here

#remove asterisks
pdClean <- pdHM %>%
  mutate(t180 = gsub("\\*", "", t180)) %>%
  mutate(t225 = gsub("\\*", "", t225))

#change these columns to numeric
pdClean$t180 <- as.integer(pdClean$t180)
pdClean$t225 <- as.integer(pdClean$t225)


```


Get mean number of prey deliveries for each study site, by time interval

```{r}

#create independet dfs for Hayman and Missouri. (since the last observation for Hayman is at t240, we'll end both datasets there)
pdHFSA <- pdClean %>%
  filter(study_site == "C") %>%
  select(t15:t240)
pdMGSA <- pdClean %>%
  filter(study_site =="B") %>%
  select(t15:t240)

#change column names to remove "t" in front of time interval
colnames(pdHFSA) <- c("15", "30", "45", "60", "75", "90", "105", "120", "135", "150", "165", "180", "195", "210", "225", "240")
colnames(pdMGSA) <- c("15", "30", "45", "60", "75", "90", "105", "120", "135", "150", "165", "180", "195", "210", "225", "240")
```




## Mean PD tables

Two stand-alone data.frames, one for HFSA and one for MGSA:

```{r}

meanHFSA <- data.frame(HGSA = colMeans(pdHFSA, na.rm = TRUE))

meanMGSA <- data.frame(MGSA = colMeans(pdMGSA, na.rm = TRUE))


```



Combined data.frame for HFSA and MGSA:

```{r}

meanPD <- data.frame(
  study_area = c(rep("HFSA", nrow(meanHFSA)), rep("MGSA", nrow(meanMGSA))),
  mean = c(colMeans(pdHFSA, na.rm = TRUE), colMeans(pdMGSA, na.rm = TRUE)),
  time = c(as.numeric(rownames(meanHFSA)), as.numeric(rownames(meanHFSA)))
  )

```



# Calculate confidence intervals


```{r}

#function to grab CIs from any vector
getCI <- function(vec, n_samp=10000) {
  smp <- replicate(n_samp, mean(sample(vec, replace = TRUE), na.rm = TRUE))
  CIs <-quantile(smp, c(0.025, 0.975), na.rm = T)
  return(CIs)
}

#now apply it across the columns of PD data
ciHFSA <- apply(pdHFSA, 2, FUN = getCI)
ciMGSA <- apply(pdMiss, 2, FUN = getCI)
```


Add these to a data frame.

```{r}

ciMeanPD <- data.frame(
  study_area = c(rep("HFSA", nrow(meanHFSA)), rep("MGSA", nrow(meanMGSA))),
  mean = c(colMeans(pdHFSA, na.rm = TRUE), colMeans(pdMGSA, na.rm = TRUE)),
  ci_l = c(ciHFSA[1,], ciMGSA[1,]),
  ci_h = c(ciHFSA[2,], ciMGSA[2,]),
  time = c(as.numeric(rownames(meanHFSA)), as.numeric(rownames(meanHFSA)))
  )

```

Plot:
```{r}

ggplot(data = ciMeanPD) +
  geom_point(aes(x = time, y = mean, color = study_area, group = study_area),
             position = position_dodge(width=0.75)) +
  geom_errorbar(aes(x=time, ymax = ci_h, ymin=ci_l, color = study_area, 
                    group = study_area),
                position = position_dodge(width=0.75)) +
  labs(x = "Time After Sunset (minutes)", y = "Mean Prey Deliveries", 
       title = "mean Prey Deliveries in HFSA and MGSA", color = "Study Area") +
  theme_minimal()

```
