---
title: "Prey Delivery Analysis"
author: "Eliza Stein"
date: "11/8/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


# Libraries

```{r}

#install.packages("tidyverse")
library(tidyverse)

```



# Load data

```{r}

setwd("C:/Users/eliza/Documents/final-project")
pdOriginal <- read.csv("./data/pd_main.csv")


#rename the first column, which imported with a special character
names(pdOriginal)[1] <- "nest"


```




## Clean up data


Filter to only include males in B (Missouri) and C (Hayman)

```{r}

pdHM <- pdOriginal %>%
  separate(col = nest, into = c("study_site", "territory"), sep = 1, remove = TRUE) %>%
  filter(study_site == "B" | study_site == "C", sex == "M")

```


Check structure of data

```{r}

str(pdHM) #columns 't180' and 't225' are character classes, "nest_age" is character class

unique(pdHM$t180) #at least one cell has an asterisk after the value
unique(pdHM$t225) #same here
unique(pdHM$nest_age) #"pred" and "" can be converted to NA

#remove asterisks
pdClean <- pdHM %>%
  mutate(t180 = gsub("\\*", "", t180)) %>%
  mutate(t225 = gsub("\\*", "", t225))


#change these columns to numeric
pdClean$t180 <- as.integer(pdClean$t180)
pdClean$t225 <- as.integer(pdClean$t225)
pdClean$nest_age <- as.integer(pdClean$nest_age)


#gather times into long form
pdGathered <- pdClean %>%
  gather(t15:t540, key = "time", value = "num_pds") %>%
   separate(
    col = time, into = c(NA, "time"), 
    sep = 1, remove = TRUE)


#make "time" numeric
pdGathered$time <- as.numeric(pdGathered$time)


```



# Prey Deliveries Throughout Night

Data was separated by study site (HFSA vs. MGSA) and by incubation vs. nestling stage. Nestling period is defined as nest_age >= 22 days. If nest age was not indicated in original dataset, field notes were used to determine whether nest was in incubation (all eggs) or nestling (at least one nestling) stage. For these records, the following values were manually input: nest_age = 0 for incubation or nest_age = 100 for nestling, so that this data could be easily separated from known nest age.  If it was later determined that nest had been predated before observation, "pred" was entered. If the nest stage could not be determined, it was left blank."pred" and "" values were converted to NA earlier when this column was converted to numeric.

since the last observation for HFSA is at time = 240, we'll end both datasets there.

First, create data frame with PDs for whole dataset, then broken down by each study site and stage.

```{r}

#create independet dfs for HFSA (nestling and incubation stage) and MGSA (nestling and incubation state). 

#select relevant columns, add column for stage, rename study sites, drop NAs
pdStage <- pdClean %>%
  select(study_site, nest_age, t15:t240) %>%
  mutate(
    stage =
      ifelse(nest_age < 22, "incubation", "nestling"),
    study_site =
      ifelse(study_site == "B", "MGSA", "HFSA")) %>%
  drop_na(stage)

#change column names to remove "t" in front of time interval
colnames(pdStage) <- c("study_site", "nest_age", "15", "30", "45", "60", "75", "90", "105", "120", "135", "150", "165", "180", "195", "210", "225", "240", "stage")

#create independent dfs for each study site and stage
pdHFSA_inc <- pdStage %>%
  filter(study_site =="HFSA", stage == "incubation") %>%
  select('15':'240')

pdHFSA_nest <- pdStage %>% 
  filter(study_site =="HFSA", stage == "nestling") %>%
  select('15':'240')

pdMGSA_inc <- pdStage %>%
  filter(study_site =="MGSA", stage == "incubation") %>%
  select('15':'240')
         
pdMGSA_nest <- pdStage %>%
  filter(study_site =="MGSA", stage == "nestling") %>%
  select('15':'240')



```




## Mean PD tables

Four stand-alone data.frames, one for (incubation), one for HFSA (nestling), one for MGSA(incubation), and one for MGSA (nestling):

```{r}

meanHFSA_inc <- data.frame(
  time = as.numeric(colnames(pdHFSA_inc)),
  HGSA_incubation = colMeans(pdHFSA_inc, na.rm = TRUE))

meanHFSA_nest <- data.frame(
  time = as.numeric(colnames(pdHFSA_nest)),
  HGSA_nestling = colMeans(pdHFSA_nest, na.rm = TRUE))

meanMGSA_inc <- data.frame(
  time = as.numeric(colnames(pdMGSA_inc)),
  MGSA_incubation = colMeans(pdMGSA_inc, na.rm = TRUE))

meanMGSA_nest <- data.frame(
  time = as.numeric(colnames(pdMGSA_nest)),
  MGSA_nestling = colMeans(pdMGSA_nest, na.rm = TRUE))

```


Two tables, one for nestling stage and one for incubation stage:

```{r}

#filtered this one to end at time = 180, because no observations exist for MGSA after that point
meanInc <- filter(
  data.frame(
    time = as.numeric(colnames(pdHFSA_inc)),
    HGSA_incubation = colMeans(pdHFSA_inc, na.rm = TRUE),
    MGSA_incubation = colMeans(pdMGSA_inc, na.rm = TRUE)),
  time <= 180
)

meanNest <-
  data.frame(
  time = as.numeric(colnames(pdHFSA_nest)),
  HGSA_nestling = colMeans(pdHFSA_nest, na.rm = TRUE),
  MGSA_nestling = colMeans(pdMGSA_nest, na.rm = TRUE)
)

```


One table showing mean PDs for all study sites and stages:
```{r}
meanAll <- data.frame(
  time = as.numeric(colnames(pdHFSA_inc)),
  HGSA_incubation = colMeans(pdHFSA_inc, na.rm = TRUE),
  HGSA_nestling = colMeans(pdHFSA_nest, na.rm = TRUE),
  MGSA_incubation = colMeans(pdMGSA_inc, na.rm = TRUE),
  MGSA_nestling = colMeans(pdMGSA_nest, na.rm = TRUE)
)

```



Combined data.frame for HFSA and MGSA:
remove?
```{r}

meanPD2 <- data.frame(
  study_area = c(rep("HFSA", nrow(meanHFSA)), rep("MGSA", nrow(meanMGSA))),
  mean = c(colMeans(pdHFSA, na.rm = TRUE), colMeans(pdMGSA, na.rm = TRUE)),
  time = c(as.numeric(rownames(meanHFSA)), as.numeric(rownames(meanHFSA)))
  )
```



## Calculate confidence intervals


```{r}

#function to grab CIs from any vector
getCI <- function(vec, n_samp=10000) {
  smp <- replicate(n_samp, mean(sample(vec, replace = TRUE), na.rm = TRUE))
  CIs <-quantile(smp, c(0.025, 0.975), na.rm = T)
  return(CIs)
}

#spread meanPD to apply function
#pdSpread <- spread(pdStage, key = "time", value = "num_pds")

#now apply it across the columns of PD data
ciHFSA_inc <- apply(pdHFSA_inc, 2, FUN = getCI)
ciHFSA_nest <- apply(pdHFSA_nest, 2, FUN = getCI)
ciMGSA_inc <- apply(pdMGSA_inc, 2, FUN = getCI)
ciMGSA_nest <- apply(pdMGSA_nest, 2, FUN = getCI)
```


Add these to a data frame. Remove rows where time > 180 for ciInc because no data is available for MGSA after this time.

```{r}

ciInc <- filter(
  data.frame(
    study_area = c(rep("HFSA", nrow(meanHFSA_inc)), rep("MGSA", nrow(meanMGSA_inc))),
    mean = c(colMeans(pdHFSA_inc, na.rm = TRUE), colMeans(pdMGSA_inc, na.rm = TRUE)),
    ci_l = c(ciHFSA_inc[1,], ciMGSA_inc[1,]),
    ci_h = c(ciHFSA_inc[2,], ciMGSA_inc[2,]),
    time = c(as.numeric(rownames(meanHFSA_inc)), as.numeric(rownames(meanMGSA_inc))),
    stage = "Incubation"),
  time <= 180)


ciNest <- data.frame(
    study_area = c(rep("HFSA", nrow(meanHFSA_nest)), rep("MGSA", nrow(meanMGSA_nest))),
    mean = c(colMeans(pdHFSA_nest, na.rm = TRUE), colMeans(pdMGSA_nest, na.rm = TRUE)),
    ci_l = c(ciHFSA_nest[1,], ciMGSA_nest[1,]),
    ci_h = c(ciHFSA_nest[2,], ciMGSA_nest[2,]),
    time = c(as.numeric(rownames(meanHFSA_nest)), as.numeric(rownames(meanMGSA_nest))),
    stage = "Nestling")
    
ciAll <- rbind(ciInc, ciNest)

```

might not need this anymore:
```{r}
ciInc <- data.frame(
  study_area = c(rep("HFSA", nrow(meanHFSA)), rep("MGSA", nrow(meanMGSA))),
  mean = c(colMeans(pdHFSA, na.rm = TRUE), colMeans(pdMGSA, na.rm = TRUE)),
  ci_l = c(ciHFSA[1,], ciMGSA[1,]),
  ci_h = c(ciHFSA[2,], ciMGSA[2,]),
  time = c(as.numeric(rownames(meanHFSA)), as.numeric(rownames(meanHFSA)))
  )

```

# Plot:

Incubation:
```{r}

plotInc <- ggplot(data = ciInc) +
  geom_point(aes(x = time, y = mean, color = study_area, group = study_area),
             position = position_dodge(width=0.75)) +
  geom_errorbar(aes(x=time, ymax = ci_h, ymin=ci_l, color = study_area, 
                    group = study_area),
                position = position_dodge(width=0.75)) +
  labs(x = "Time After Sunset (minutes)", y = "Mean Prey Deliveries", 
       title = "Incubation", color = "Study Area") +
  theme_minimal()

```


Nestling:
```{r}

plotNest <- ggplot(data = ciNest) +
  geom_point(aes(x = time, y = mean, color = study_area, group = study_area),
             position = position_dodge(width=0.75)) +
  geom_errorbar(aes(x=time, ymax = ci_h, ymin=ci_l, color = study_area, 
                    group = study_area),
                position = position_dodge(width=0.75)) +
  labs(x = "Time After Sunset (minutes)", y = "Mean Prey Deliveries", 
       title = "Nestling", color = "Study Area") +
  theme_minimal()

```

Both:
```{r}

ggplot(data = ciAll) +
  geom_point(aes(x = time, y = mean, color = study_area, group = study_area),
             position = position_dodge(width=0.75)) +
  geom_errorbar(aes(x=time, ymax = ci_h, ymin=ci_l, color = study_area, 
                    group = study_area),
                position = position_dodge(width=0.75)) +
  labs(x = "Time After Sunset (minutes)", y = "Mean Prey Deliveries", 
       title = "Average Prey Deliveries Throughout Night", color = "Study Area") +
  theme_minimal() +
  facet_grid(stage ~ .)

```

