---
title: "Prey Delivery Analysis"
author: "Eliza Stein"
date: "11/8/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
break down by nestling/incubation
for Hayman nestling and incubation: dot and whisker plot, if pd changes over years

# Libraries

```{r}

#install.packages("tidyverse")
library(tidyverse)

```



# Load data

```{r}

setwd("C:/Users/eliza/Documents/final-project")
pdOriginal <- read.csv("./data/pd_main.csv")


#rename the first column, which imported with a special character
names(pdOriginal)[1] <- "nest"


```




## Clean up data


Filter to only include males in B (Missouri) and C (Hayman)

```{r}

pdHM <- pdOriginal %>%
  separate(col = nest, into = c("study_site", "territory"), sep = 1, remove = TRUE) %>%
  filter(study_site == "B" | study_site == "C", sex == "M")

```


Check structure of data

```{r}

str(pdHM) #columns 't180' and 't225' are character classes, "nest_age" is character class

unique(pdHM$t180) #at least one cell has an asterisk after the value
unique(pdHM$t225) #same here
unique(pdHM$nest_age) #"pred" and "" can be converted to NA

#remove asterisks
pdClean <- pdHM %>%
  mutate(t180 = gsub("\\*", "", t180)) %>%
  mutate(t225 = gsub("\\*", "", t225))


#change these columns to numeric
pdClean$t180 <- as.integer(pdClean$t180)
pdClean$t225 <- as.integer(pdClean$t225)
pdClean$nest_age <- as.integer(pdClean$nest_age)


#gather times into long form
pdGathered <- pdClean %>%
  gather(t15:t540, key = "time", value = "num_pds") %>%
   separate(
    col = time, into = c(NA, "time"), 
    sep = 1, remove = TRUE)


#make "time" numeric
pdGathered$time <- as.numeric(pdGathered$time)


```



# Prey Deliveries Throughout Night

Data was separated by study site (HFSA vs. MGSA) and by incubation vs. nestling stage. Nestling period is defined as nest_age >= 22 days. If nest age was not indicated in original dataset, field notes were used to determine whether nest was in incubation (all eggs) or nestling (at least one nestling) stage. For these records, the following values were manually input: nest_age = 0 for incubation or nest_age = 100 for nestling, so that this data could be easily separated from known nest age.  If it was later determined that nest had been predated before observation, "pred" was entered. If the nest stage could not be determined, it was left blank."pred" and "" values were converted to NA earlier when this column was converted to numeric.

Get mean number of prey deliveries for each study site, by time interval:

```{r}

#create independet dfs for HFSA (nestling and incubation stage) and MGSA (nestling and incubation state). (since the last observation for Hayman is at t240, we'll end both datasets there)
pdIncH <- pdClean %>%
  filter(study_site == "C", nest_age < 22) %>%
  select(study_site, nest_age, t15:t240)


pdNestH <- pdClean %>%
  filter(study_site == "C", nest_age >= 22) %>%
  select(study_site, nest_age, t15:t240)

#winner: select relavent columns, add column for stage, rename study sites, drop NAs
pdStage <- pdGathered %>%
  select(study_site, nest_age, time, num_pds) %>%
  mutate(
    stage =
      ifelse(nest_age < 22, "incubation", "nestling"),
    study_site =
      ifelse(study_site == "B", "MGSA", "HFSA")) %>%
  drop_na(num_pds, stage)
  
      
mutate(t180 = gsub("\\*", "", t180))

pdMGSA <- pdClean %>%
  filter(study_site =="B") %>%
  select(t15:t240)

#change column names to remove "t" in front of time interval
colnames(pdHFSA) <- c("15", "30", "45", "60", "75", "90", "105", "120", "135", "150", "165", "180", "195", "210", "225", "240")
colnames(pdMGSA) <- c("15", "30", "45", "60", "75", "90", "105", "120", "135", "150", "165", "180", "195", "210", "225", "240")
```




## Mean PD tables

Two stand-alone data.frames, one for HFSA and one for MGSA:

#Maybe take this out:
```{r}

meanHFSA <- data.frame(HGSA = colMeans(pdHFSA, na.rm = TRUE))

meanMGSA <- data.frame(MGSA = colMeans(pdMGSA, na.rm = TRUE))


```



Combined data.frame for HFSA and MGSA:

```{r}

meanPD <- data.frame(
  study_area = c(rep("HFSA", nrow(meanHFSA)), rep("MGSA", nrow(meanMGSA))),
  mean = c(colMeans(pdHFSA, na.rm = TRUE), colMeans(pdMGSA, na.rm = TRUE)),
  time = c(as.numeric(rownames(meanHFSA)), as.numeric(rownames(meanHFSA)))
  )

# winner: get mean for each study site by time and nest stage, filter out rows >225 bc HGSA has no more values
meanPD <- pdStage %>%
  group_by(time, study_site, stage) %>%
  summarise(mean = mean(num_pds, na.rm = TRUE)) %>%
  filter(time <= 225)

  

```



## Calculate confidence intervals


```{r}

#function to grab CIs from any vector
getCI <- function(vec, n_samp=10000) {
  smp <- replicate(n_samp, mean(sample(vec, replace = TRUE), na.rm = TRUE))
  CIs <-quantile(smp, c(0.025, 0.975), na.rm = T)
  return(CIs)
}

#spread meanPD to apply function
pdSpread <- spread(pdStage, key = "time", value = "num_pds")

#now apply it across the columns of PD data
ciHFSA <- apply(pdSpread$'15':pdSpread$'225', 2, FUN = getCI)
ciMGSA <- apply(pdSpread, 2, FUN = getCI)
```


Add these to a data frame.

```{r}

ciMeanPD <- data.frame(
  study_area = c(rep("HFSA", nrow(meanHFSA)), rep("MGSA", nrow(meanMGSA))),
  mean = c(colMeans(pdHFSA, na.rm = TRUE), colMeans(pdMGSA, na.rm = TRUE)),
  ci_l = c(ciHFSA[1,], ciMGSA[1,]),
  ci_h = c(ciHFSA[2,], ciMGSA[2,]),
  time = c(as.numeric(rownames(meanHFSA)), as.numeric(rownames(meanHFSA)))
  )

```

Plot:
```{r}

ggplot(data = ciMeanPD) +
  geom_point(aes(x = time, y = mean, color = study_area, group = study_area),
             position = position_dodge(width=0.75)) +
  geom_errorbar(aes(x=time, ymax = ci_h, ymin=ci_l, color = study_area, 
                    group = study_area),
                position = position_dodge(width=0.75)) +
  labs(x = "Time After Sunset (minutes)", y = "Mean Prey Deliveries", 
       title = "Mean Prey Deliveries in HFSA and MGSA", color = "Study Area") +
  theme_minimal()

```
